Expert-In-The-Loop Causal Discovery: \\ Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
Minor changes

\section{Background}
\label{sec:background}
We denote random variables using uppercase letters, such as $ X $, and a set of
random variables by $ \bm{X} = \{X_1, \cdots, X_k\} $, where $ \rvert \bm{X}
\rvert = k $ indicates the number of variables in the set. We denote the
covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, the
covariance matrix as $ \Sigma $, and the entry corresponding to covariance
between $ X $ and $ Y $ using $ \Sigma_{XY} $. In the context of this paper, we
consider causal DAGs that contain only observed variables. The structure of the
DAG is represented using a graph $ G = (V, E) $, where $ V $ are the nodes
corresponding to the random variables, and $ E $ are the edges denoting the
causal relationship between the variables. The set of parents of $ X $ in $ G $
are denoted as $ \textrm{Pa}_G(X) $, and its ancestors and descendants as $
\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $ respectively. We denote a path, $
\pi(X, Y) = \{ X, V_0, V_1, \cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G
$ where each consecutive pair of variables in $ \pi $ are connected through an
edge. We say $ X $ and $ Y $ are d-connected in $ G $ given a conditioning set
$ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) for every
collider structure ($ V_1 \rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $
\bm{Z} \cap \{ V_2, \textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every
non-collider structure ($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1
\leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
d-separated given $ \bm{Z} $ if no d-connecting path exist between them.

>>>>

\section{Background}
\label{sec:background}
We denote random variables using uppercase letters, such as $ X $, and a set of
random variables by $ \bm{X} = \{X_1, \cdots, X_k\} $, where $ \rvert \bm{X}
\rvert = k $ indicates the number of variables in the set. We denote the
covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, the
covariance matrix as $ \Sigma $, and the entry corresponding to the covariance
between $ X $ and $ Y $ using $ \Sigma_{XY} $. In this paper, we
consider causal DAGs that contain only observed variables. The structure of the
DAG is represented using a graph $ G = (V, E) $, where $ V $ are the nodes
corresponding to the random variables, and $ E $ are the edges denoting the
causal relationship between the variables. The set of \emph{parents} of $ X $ in $ G $
are denoted as $ \textrm{Pa}_G(X) $, and its \emph{ancestors} and \emph{descendants} as $
\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $ respectively. We denote a \emph{path}, $
\pi(X, Y) = \{ X, V_0, V_1, \cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G
$ where each consecutive pair of variables in $ \pi $ are connected through an
edge. We say $ X $ and $ Y $ are \emph{d-connected} in $ G $ given a conditioning set
$ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) for every
collider structure ($ V_1 \rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $
\bm{Z} \cap \{ V_2, \textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every
non-collider structure ($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1
\leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
\emph{d-separated} given $ \bm{Z} $ if no d-connecting path exist between them.
